Heuristics and Really Hard Instances for Subgraph Isomorphism Problems

### 2016-01-14 paper/paper.tex
Betterer writing

We create a pseudo-boolean (PB) encoding as follows. For each pattern vertex $v$ and each target
vertex $w$, create a binary variable which takes the value 1 if and only if that $v$ is mapped to
$w$.  Constraints are added to ensure that each pattern vertex maps to exactly one target vertex,
that each target vertex is mapped to by at most one pattern vertex, and that adjacent edges are
mapped to adjacent edges. Because we are solving the induced problem, we also add constraints to
ensure that non-adjacent edges are mapped to non-adjacent edges. We used the Clasp
solver~\citep{gekakaosscsc11a} to solve the pseudo-boolean instances. The instances that are hard
for the Glasgow solver remain hard for the PB solver, including instances inside the central region,
and the easy satisfiable instances remain easy.

@@

?? Clique encoding: association graph, a version of BBMC \citep{SanSegundo:2011} modified to solve the
decision problem. The hard instances in the central region remain hard. Additionally, some of the
easy satisfiable instances become hard. Possible weakness in variable ordering heuristics. The
interesting thing is here that clique knows nothing about adjacency or non-adjacency.

?? Some easy instances become hard, here, but hard instances remain hard. In particular, the large
central region for the larger pattern graphs which is not near the phase transitions is still
relatively difficult in all encodings. This suggests that these instances may genuinely be hard,
despite not being near a phase transition, and that this does not indicate a weakness in subgraph
isomorphism solvers.

>>>>

We create a pseudo-boolean (PB) encoding as follows. For each pattern vertex $v$ and each target
vertex $w$, we have a binary variable which takes the value 1 if and only if that $v$ is mapped to
$w$.  Constraints are added to ensure that each pattern vertex maps to exactly one target vertex,
that each target vertex is mapped to by at most one pattern vertex, that adjacent vertices are
mapped to adjacent vertices, and that non-adjacent vertices are mapped to non-adjacent vertices. We
used the Clasp solver~\citep{gekakaosscsc11a} to solve the pseudo-boolean instances. The instances
that are hard for the Glasgow solver remain hard for the PB solver, including instances inside the
central region, and the easy satisfiable instances remain easy.

@@

We also reduced the problem to the clique problem via the association graph encoding show in
\cref{figure:sip}, and solved the instances using a version of BBMC \citep{SanSegundo:2011} modified
to solve the decision problem. Again, the instances in the central region remain hard, and
additionally, some of the easy satisfiable instances become hard.

These results suggest that the central region may be genuinely hard, despite not being near a phase
transition.  The clique encoding in particular rules out the hypothesis that subgraph isomorphism
solvers only find this region hard due to not reasoning simultaneously about adjacency and
non-adjacency, since the association graph encoding constraints consider compatibility rather than
adjacency and non-adjacency.
